### 2016-07-29 QPLIB.tex
affiliation updated (please check)

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  QPLIB.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\newif\ifMPC
\global\MPCtrue

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
 \documentclass[smallextended]{svjour3}
 \smartqed
 \journalname{Mathematical Programming Computation}
\else
 \documentclass[12pt,A4wide]{article}
\fi

\usepackage{url}
\usepackage{secdot}
\usepackage{color}
\usepackage{booktabs}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{todonotes}

\usepackage{longtable}
\usepackage{enumitem}
\usepackage{calc}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% page size

\ifMPC
\else
 \setlength{\textwidth}{16 cm}
 \setlength{\topmargin}{0 cm}
 \setlength{\headsep}{0 cm}
 \setlength{\textheight}{24 cm}
 \setlength{\evensidemargin}{0cm}
 \setlength{\oddsidemargin}{0cm}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\def\NN{\mathbb{N}}
\def\RR{\mathbb{R}}
\def\Mcq{\mathcal{Q}}
\def\Mcl{\mathcal{L}}
\def\Mcm{\mathcal{M}}
\def\Mcn{\mathcal{N}}
\def\Mcb{\mathcal{B}}
\def\Mcr{\mathcal{R}}
\def\Mcu{\mathcal{U}}
\def\Mcx{\mathcal{X}}
\def\Mcz{\mathcal{Z}}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% Theorem-like environment definitions

\ifMPC
\else
\newtheorem{theorem}{Theorem}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}{Remark}

\newenvironment{example}{\begin{EX}\rm}{\end{EX}}

\newenvironment{proof}[1][Proof]{\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
%\renewenvironment{proof}{{\bfseries Proof}}{\qed}

\def\remark{\stepcounter{theorem}\par{\bf Remark}.
\ignorespaces}
\def\endremark{ \vbox{\hrule\hbox{%
    \vrule height1.3ex\hskip0.8ex\vrule}\hrule
   }\par}

\newenvironment{keywords}{\vspace{0.4in}\begin{quote}\small \em
{\bf Keywords\/}:}{\end{quote}}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\begin{document}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\title{QPLIB: A Library of Quadratic Programming Instances}

\ifMPC
\titlerunning{QPLIB}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
\author{Fabio Furini \and
        Emiliano Traversi \and
        Pietro Belotti \and
        Antonio Frangioni \and
        Ambros Gleixner \and
        Nick Gould \and
        Leo Liberti \and
        Andrea Lodi \and
        Ruth Misener \and
        Hans Mittelmann \and
        Nick Sahinidis \and
        Stefan Vigerske \and
        Angelika Wiegele
        }

\institute{Fabio Furini \at
           LAMSADE, Universit{\'e} Paris Dauphine, 75775 Paris, France.
           \email{fabio.furini@dauphine.fr}
%
\and
           Emiliano Traversi \at
           LIPN, Universit{\'e} de Paris 13, 93430 Villetaneuse, France.
           \email{emiliano.traversi@lipn.fr}
%
\and       Pietro Belotti \at
           Xpress-Optimizer team, FICO, Birmingham UK
           \email{pietrobelotti@fico.com}
%
\and       Antonio Frangioni \at
           Dipartimento di Informatica, Universit\`a di Pisa,
           Largo B. Pontecorvo 2, 56127 Pisa, Italy,
           \email{frangio@di.unipi.it}
%
\and       Ambros Gleixner \at
           Zuse Institute Berlin, Takustr. 7, 14195 Berlin, Germany.
           \email{gleixner@zib.de}
%
\and       Nick Gould \at
           STFC-Rutherford Appleton Laboratory, Chilton, Oxfordshire OX11 0QX,
           England.
           \email{nick.gould@stfc.ac.uk}
%
\and       Leo Liberti \at
           CNRS LIX, Ecole Polytechnique, 91128 Palaiseau, France.
           \email{liberti@lix.polytechnique.fr}
%
\and       Andrea Lodi \at
           Ecole Polytechnique de Montréal, Canada.               \email{andrea.lodi@polymtl.ca}
%
\and       Ruth Misener \at
           Department of Computing,Imperial College London, UK. \email{r.misener@imperial.ac.uk}
%
\and       Hans Mittelmann \at
           School of Mathematical and Statistical Sciences, Arizona State University,      Tempe, AZ 85287-1804, U.S.A.
\email{mittelmann@asu.edu}
%
\and       Nick Sahinidis \at
Chemical Engineering, Carnegie Mellon University, U.S.A. \email{sahinidis@cmu.edu}
%
\and       Stefan Vigerske \at
           GAMS Software GmbH, P.O. Box 40 59,50216 Frechen, Germany.
           \email{stefan@gams.com}
%
\and       Angelika Wiegele \at
Institut für Mathematik, Alpen-Adria-Universität Klagenfurt, Austria.  \email{angelika.wiegele@aau.at}
           }

\else

\author{Fabio Furini\\
%
\and
        Emiliano Traversi\\
%
\and    Pietro Belotti\\
%
\and    Antonio Frangioni\\
%
\and    Ambros Gleixner\\
%
\and    Nick Gould\\
%
\and    Leo Liberti\\
%
\and    Andrea Lodi\\
%
\and    Ruth Misener\\
%
\and    Hans Mittelmann\\
%
\and    Nick Sahinidis\\
%
\and    Stefan Vigerske\\
%
\and    Angelika Wiegele
        }
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\ifMPC
 \date{Received: date / Accepted: date}
\else
 \date{}
\fi

\maketitle

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP).
 QP is a very ``varied'' class, comprising sub-classes of problems ranging from
 trivial to undecidable. Solution methods for QP are very diverse, ranging from
 entirely combinatorial ones to completely continuous ones, including many for
 which both aspects are fundamental. Selecting a set of instances of QP that is
 at the same time not overwhelmingly numerous and sufficiently challenging for the
 many different interested communities is therefore important. We propose a simple
 taxonomy for QP instances that leads to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and
 solvers. Finally we describe how the library was put together, and detail its final
 contents.
\end{abstract}

\ifMPC
 \keywords{Instance Library, Quadratic Programming}

 \subclass{90C06 \and 90C25}
\else
 \begin{keywords}
  Instance Library, Quadratic Programming
 \end{keywords}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-1.0.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.1.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.2.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.3.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-2.4.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-3.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{QPLIB-4.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\input{conclusion.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\section{Acknowledgements}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\bibliography{biblio}

\ifMPC
 \bibliographystyle{spmpsci}
\else
 \bibliographystyle{plain}
\fi

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\appendix

\input{Appendix.tex}

%  QPLIB format. Move to main text?

\input{format.tex}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\end{document}

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%  End QPLIB.tex
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -