### 2017-01-24 conclusion.tex
revise conclusions and appendix

\section{Conclusions}\label{sec:conclusions}

 This manuscript describes the first comprehensive library of instances for Quadratic Programming (QP). Since QP comprises  different and ``varied'' categories of problems, we propose a classification and then we describe the principal solution methods representing the state-of-the-art algorithms for QP.

We then describe the steps of the adopted process used to filter the gathered instances  in order to build the new library. The goals we purposed are of different natures and we tried to build a library which is computationally challenging and as broad as possible, i.e., it represents the largest possible categories of QP problems.

We want to stress once again that we intentionally avoid to perform a computational comparison of the performances of the different solvers. Our goal is instead to provide a common test-bed of instances for practitioners and researchers in the field. This new library will hopefully serve as a point of reference to test new ideas and algorithms for QP problems.